ITHIM

### 2021-10-18 latex/output-md.Rmd
Fix layout and typos

@@

![ITHIM workflow. Inputs to the model have a bold, black outline. The
graph depicts parent--child relationships, where a child (at the head of
an arrow) depends upon all its parents (at the source(s) of the
arrow(s)). To aid visualisation, an arrow connected to the outside of a
grey box indicates that the arrow connects to each item within the box.
*AT: active travel. MET: metabolic equivalent task. pp: per person. AP:
air pollution. PA: physical activity. DR: dose--response relationship.
GBD: global burden of disease. BOD: (local) burden of
disease.*](flowchart.pdf){#workflow width="90%"}

@@

::: {.center}
::: {#indices}
  Label   Name                   Levels
  ------- ---------------------- ---------------------
  $a$     Age group              15--49
                                 50--69
  $g$     Gender                 Male
                                 Female
  $h$     Disease                
  $i$     Individual index       
  $j$     Trip index             
  $m$     Transport mode         walk
                                 cycling (cyc.)
                                 bus
                                 car
                                 motorbike (mot.)
                                 goods vehicles (GV)
                                 subway (sub.)
  $o$     Outcome                Death
                                 YLL
  $s$     Scenario               
  $w$     AP RR hyperparameter   $1,2,3,4$
  $x$     PA RR hyperparameter   $1,2,3$
  $z$     PA dose                $0,...,300$

@@

::: {.center}
::: {#abbrev}
  Abbreviation   Meaning
  -------------- -----------------------------------------------
  AP             Air pollution
  CDF            Cumulative distribution function
  COPD           Chronic obstructive pulmonary disease
  DR             Dose response
  EVPPI          Expected value of perfect partial information
  GAM            Generalised additive model
  GBD            Global burden of disease
  IHD            Ischemic heart disease
  ITHIM          Integrated transport and health impact model
  LRI            Lower respiratory infection
  MMET           Marginal metabolic equivalent task
  PA             Physical activity
  PAF            Population attributable fraction
  PIF            Potential impact fraction
  PM2.5          Particulate matter (diameter $< 2.5 \mu$m)
  PT             Public transport
  RR             Relative risk
  VOI            Value of information
  YLL            Years of life lost

@@

::: {.landscape}
::: {.center}

@@

  ------------------------------------------- ----------------------------------------------------------------- ---------------------------------- ------------------------------- ----------------
  [\[inputs\]]{#inputs label="inputs"}                                                                          *ITHIM-R reference*                                                
  Name                                        Description                                                       Value in Accra model                                         Label   Model/ Setting
  *Global values*                                                                                                                                                                  
  $U_{a,g,h,o}$                               Background burden of disease                                      Constant                                                  GBD_DATA          Setting
  $N_{a,g}$                                   Population by age and gender in GBD_DATA                          Constant                                                  GBD_DATA          Setting
  $\bar{N}_{a,g}$                             Population by age and gender                                      Constant                                               DEMOGRAPHIC          Setting
  $\rho_h$                                    Chronic disease scalar                                            1 or Lnorm(0,0.18)                          CHRONIC_DISEASE_SCALAR          Setting
  *Travel values*                                                                                                                                                                  
  $Z_{m=walk}$                                Walk speed                                                        4.8                                                    MODE_SPEEDS          Setting
  $Z_{m=cyc.}$                                Cycle speed                                                       14.5                                                   MODE_SPEEDS          Setting
  $Z_{m=bus}$                                 Bus speed                                                         15                                                     MODE_SPEEDS          Setting
  $Z_{m=car}$                                 Car speed                                                         21                                                     MODE_SPEEDS          Setting
  $Z_{m=mot.}$                                Motorbike speed                                                   25                                                     MODE_SPEEDS          Setting
  $\epsilon$                                  Walk time to bus                                                  5 or Lnorm(log(5),0.18)                              BUS_WALK_TIME          Setting
  $\mu_{m=HGV}$                               Truck distance relative to car                                    0.21 or Beta(3,10)                              TRUCK_TO_CAR_RATIO          Setting
  $\mu_{m=bus}$                               Bus driver distance relative to bus passenger distance            1/45 or Beta(20,600)                        BUS_TO_PASSENGER_RATIO          Setting
  $\mu_{m=mot.}$                              Motorcycle fleet distance relative to total motorcycle distance   37/68 or Beta(37/3+1,31/3+1)             FLEET_TO_MOTORCYCLE_RATIO          Setting
  $T_j$                                       Trip-level data                                                   Constant                                           SYNTHETIC_TRIPS          Setting
  *Injury model values*                                                                                                                                                            
  $I_{a,g,m_{\text{vic}},m_{\text{str}}}$     Fatalities table                                                  Constant                                              INJURY_TABLE          Setting
  $\sigma$                                    Injury reporting rate                                             1 or Beta(8,3)                               INJURY_REPORTING_RATE          Setting
  $\omega$                                    Sum of distance exponents                                         1.9 or Lnorm(log(1.9),log(1.03))                  SIN_EXPONENT_SUM            Model
  $\psi$                                      Casualty fraction of $\omega$                                     0.5 or Beta(20,20)                    CASUALTY_EXPONENT\_ FRACTION            Model
  *Pollution model values*                                                                                                                                                         
  $\eta$                                      Background PM2.5 concentration                                    50 or Lnorm(log(50),0.18)                             PM_CONC_BASE          Setting
  $\zeta$                                     Fraction of measured PM2.5 concentration due to road transport    0.225 or Beta(5,20)                                 PM_TRANS_SHARE          Setting
  $\gamma_{m}$                                Vehicle emission inventory                                        Constant or Dirichlet                           EMISSION_INVENTORY          Setting
  $P$                                         Vehicle emission confidence                                       1                                    EMISSION_INVENTORY_CONFIDENCE          Setting
  *Pollution & health model values*                                                                                                                                                
  $C_1$                                       Base-level inhalation rate                                        1                                     BASE_LEVEL_INHALATION \_RATE            Model
  $C_2$                                       Exposure ratio window closed                                      0.5                                       CLOSED_WINDOW_PM\_ RATIO            Model
  $C_3$                                       Proportion of travel with closed windows                          0.5                                            CLOSED_WINDOW_RATIO          Setting
  $C_4$                                       Parameter for on-road PM2.5 level                                 3.216                                               ROAD_RATIO_MAX            Model
  $C_5$                                       Parameter for on-road PM2.5 level                                 0.379                                             ROAD_RATIO_SLOPE            Model
  $C_6$                                       Exposure ratio for underground                                    0.8                                                SUBWAY_PM_RATIO          Setting
  $H_{a,h,w}$                                 AP dose--response--curve parameters                               Constant                                                     DR_AP            Model
  $\xi_{h,w}$                                 Quantiles for AP RR functions                                     0.5 or Unif(0,1)                       AP_DOSE_RESPONSE\_ QUANTILE            Model
  *Physical activity model values*                                                                                                                                                 
  $\chi$                                      Day-to-week travel scalar                                         7 or 7$\times$Beta(20,3)               DAY_TO_WEEK_TRAVEL\_ SCALAR            Model
  $Y_i$                                       Individual non-transport MMETs                                    Constant                                PA_SET\$work_ltpa_marg_met          Setting
  $\theta$                                    Background PA scalar                                              1 or Lnorm(0,0.18)                            BACKGROUND_PA_SCALAR          Setting
  $\tilde{Y}$                                 Confidence in PA survey                                           1                                         BACKGROUND_PA_CONFIDENCE          Setting
  $\tilde{\theta}$                            Background PA zeros quantile                                      0.5 or Unif(0,1)                               BACKGROUND_PA_ZEROS          Setting
  $\lambda_{m=\text{cyc.}}$                   Cycling MET surplus                                               4.63 or Lnorm(log(4.63),1)                            MMET_CYCLING            Model
  $\lambda_{m=\text{walk}}$                   Walking MET surplus                                               2.53 or Lnorm(log(2.53),1)                            MMET_WALKING            Model
  $\lambda_{m\not\in\text{walk,cyc.}}$        In-vehicle MET surplus                                            0                                                                             Model
  *Physical activity & health model values*                                                                                                                                        
  $G_{h,x,z}$                                 Look-up table for relative risk truncated normal distribution     Constant                                                                      Model
  $\phi_{h}$                                  Quantiles for PA RR functions                                     0.5 or Unif(0,1)                       PA_DOSE_RESPONSE\_ QUANTILE            Model
  ------------------------------------------- ----------------------------------------------------------------- ---------------------------------- ------------------------------- ----------------

@@

![Example of the dose--response workflow for PA and total cancer. Top
left: results of the metaanalysis: mean, lower bound and upper bound of
the relationship between PA dose and relative risk of disease. Top
right: examples of 500 samples from the normal distribution defined by
the mean of the top-left panel and the standard deviation defined by
(upper bound minus lower bound) divided by 1.96 and truncated at 0.
Bottom left: relative risks of individuals with PA doses 5, 25 and 45
are found by mapping their doses onto the median curve. This is used in
the "constant" ITHIM use case. Bottom right: relative risks of
individuals with PA doses 5, 25 and 45 for three different samples from
the distribution shown in the top-right panel. For each sample, each
individual is mapped onto the response curve it defines.
](cancer_PA_dose_response.pdf){#cancer width="60%"}

@@

The XX uncertain univariate parameters are shown in Figure
[3](#distributions){reference-type="ref" reference="distributions"}.
Most of the parameters are self explanatory. Cycling and walking MMETs
are the number of MMETs per hour when undertaking cycling and walking,
and determine also the ventilation rates. Motorcycle distance is the
total distance travelled by motorcycles relative to the total distance
travelled by cars in the baseline scenario. Non-travel PA, injury
reporting rate and NCD burden all act as scalars for the corresponding
datasets. Note that the non-travel PA scalar does not affect the
$\approx$`<!-- -->`{=html}40% of the population whose non-travel PA is
0.

![Distributions of univariate uncertain parameters (1024
samples).](uncertain_parameters.pdf){#distributions width="95%"}

@@

![Distributions of a Beta-distributed propensity to non-travel PA with
varying confidence values in the raw input data. The sampled value gives
the fraction of the population that engages in non-travel PA.
](prob_zero_PA.pdf){#prob_zero_PA width="50%"}

![Distributions of Dirichlet-distributed emission inventories for four
modes and varying confidence values. Raw input values were 4, 4, 20, and
60. ](emission_dist.pdf){#emission_dist width="65%"}

@@

![We use the same uniform samples to generate random variables from
different distributions. It means that the samples are perfectly
correlated.](DR_transform.pdf){#cdf width="80%"}

@@

# Results

@@

![YLLs in the scenario minus YLLs in the baseline for
Accra.](accra_walk_yll.pdf){#results width="80%"}

@@

![EVPPI for Accra's "walking" scenario for all causes of YLLs excluding
"all cause" and "neoplasm".](accra_walk_evppi.pdf){#evppi width="70%"}

::: {.appendix}
# ITHIM with parameters

![ITHIM-R workflow, with variable parameters in purple. Fixed inputs to
the model have a bold, black outline, which is solid for inputs we
consider "global" and dashed for inputs we consider "local". There are
four global inputs, which will be embedded in ITHIM-R. There are seven
local inputs, which users should provide. The graph depicts
parent--child relationships, where a child (at the head of an arrow)
depends upon all its parents (at the source(s) of the arrow(s)). To aid
visualisation, an arrow connected to the outside of a grey box indicates
that the arrow connects to each item within the box. *AT: active travel.
MET: metabolic equivalent task. pp: per person. AP: air pollution. PA:
physical activity. DR: dose--response relationship. GBD: global burden
of disease.* NB: $\rho$ only impacts the NCD burden of GBD, not all of
it. Missing parameter: motorcycle distance. $\lambda$ is a 2D parameter.
$\xi$ is 4D.](ithim_with_parameters.pdf){#ithim_with_parameters
width="\\textwidth"}

@@

![1000 samples from PA dose--response
curves](PA_dose_response.pdf){#PADR width="90%"}

  ------------------------------------------------------------------------------------ ---------------------------------------------------------------------------------------
   ![100 samples from AP dose--response curves](cvd_ihdDR-sim.pdf){#APDR width="40%"}   ![100 samples from AP dose--response curves](cvd_strokeDR-sim.pdf){#APDR width="40%"}
  ------------------------------------------------------------------------------------ ---------------------------------------------------------------------------------------

  ------------------------------------------------------------------------------------- -------------------------------------------------------------------------------- --------------------------------------------------------------------------------------
   ![100 samples from AP dose--response curves](neo_lungDR-sim.pdf){#APDR width="30%"}   ![100 samples from AP dose--response curves](lriDR-sim.pdf){#APDR width="30%"}   ![100 samples from AP dose--response curves](resp_copdDR-sim.pdf){#APDR width="30%"}
  ------------------------------------------------------------------------------------- -------------------------------------------------------------------------------- --------------------------------------------------------------------------------------

::: {.landscape}
# Tabulated ITHIM equations {#tableequations}

::: {.center}
::: {.ThreePartTable}
::: {.TableNotes}
CDF: cumulative distribution function. We use generative distributions
to generate random numbers so that samples are correlated. For
$t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$
with CDF CDF$_F$. (I'd like to find an improved notation for this.)
:::

@@

  Name                                                                                                                                                                                    Description                                                                                      Label
  --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- ------------------------------------------------------------------------------------------------ -------------------------------------------
  *Scenario generation*                                                                                                                                                                                                                                                                    
  \...                                                                                                                                                                                                                                                                                     
  *Travel calculations*                                                                                                                                                                                                                                                                    
  $\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk}\}\, \forall\, T_j:m(j)=\text{bus}$                                                                                           New walk trips to bus                                                                            walk_to_bus_and_combine_scen
  $\bar{T}_{a,g,m,s}=\sum_{j:a(j)=a,g(j)=g,m(j)=m,s(j)=s}Q_{i(j)}T_j$                                                                                                                     (Weighted) total travel per age, gender, mode and scenario                                       
  $\check{T}_{a,g,m,s}=\frac{\bar{T}_{a,g,m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}$                                                                                                             Relative travel per mode and scenario by age and gender                                          inj_distances\[\[1\]\]

>>>>

---
title: "ITHIM"
author: "Rob Johnson, Ali Abbas"
date: "`r format(Sys.time(), '%d %B, %Y')`"

output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

@@

!\[ITHIM workflow. Inputs to the model have a bold, black outline. The graph depicts parent--child relationships, where a child (at the head of an arrow) depends upon all its parents (at the source(s) of the arrow(s)). To aid visualisation, an arrow connected to the outside of a grey box indicates that the arrow connects to each item within the box. \*AT: active travel. MET: metabolic equivalent task. pp: per person. AP: air pollution. PA: physical activity. DR: dose--response relationship. GBD: global burden of disease. BOD:

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("flowchart.pdf")
```

@@

::: center
::: {#indices}
| Label | Name                 | Levels              |
|:------|:---------------------|:--------------------|
| $a$   | Age group            | 15--49              |
|       |                      | 50--69              |
| $g$   | Gender               | Male                |
|       |                      | Female              |
| $h$   | Disease              |                     |
| $i$   | Individual index     |                     |
| $j$   | Trip index           |                     |
| $m$   | Transport mode       | walk                |
|       |                      | cycling (cyc.)      |
|       |                      | bus                 |
|       |                      | car                 |
|       |                      | motorbike (mot.)    |
|       |                      | goods vehicles (GV) |
|       |                      | subway (sub.)       |
| $o$   | Outcome              | Death               |
|       |                      | YLL                 |
| $s$   | Scenario             |                     |
| $w$   | AP RR hyperparameter | $1,2,3,4$           |
| $x$   | PA RR hyperparameter | $1,2,3$             |
| $z$   | PA dose              | $0,...,300$         |

@@

::: center
::: {#abbrev}
| Abbreviation | Meaning                                       |
|:-------------|:----------------------------------------------|
| AP           | Air pollution                                 |
| CDF          | Cumulative distribution function              |
| COPD         | Chronic obstructive pulmonary disease         |
| DR           | Dose response                                 |
| EVPPI        | Expected value of perfect partial information |
| GAM          | Generalised additive model                    |
| GBD          | Global burden of disease                      |
| IHD          | Ischemic heart disease                        |
| ITHIM        | Integrated transport and health impact model  |
| LRI          | Lower respiratory infection                   |
| MMET         | Marginal metabolic equivalent task            |
| PA           | Physical activity                             |
| PAF          | Population attributable fraction              |
| PIF          | Potential impact fraction                     |
| PM2.5        | Particulate matter (diameter $< 2.5 \mu$m)    |
| PT           | Public transport                              |
| RR           | Relative risk                                 |
| VOI          | Value of information                          |
| YLL          | Years of life lost                            |

@@

::: landscape
::: center

@@

+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| \[\[inputs\]\]{label= "inputs"}           |                                                                 | *ITHIM-R reference*                |                                |                |
+:==========================================+=================================================================+:===================================+================================+================+
| Name                                      | Description                                                     | Value in Accra model               | Label                          | Model/ Setting |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| *Global values*                           |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $U_ { a,g,h,o}$                           | Background burden of disease                                    | Constant                           | GBD_DATA                       | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $N_{a,g}$                                 | Population by age and gender in GBD_DATA                        | Constant                           | GBD_DATA                       | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\bar { N}_{a,g}$                         | Population by age and gender                                    | Constant                           | DEMOGRAPHIC                    | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\rho_h$                                  | Chronic disease scalar                                          | 1 or Lnorm(0,0.18)                 | CHRONIC_DISEASE_SCALAR         | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| *Travel values*                           |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $Z _ {m=walk}$                            | Walk speed                                                      | 4.8                                | MODE_SPEEDS                    | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $Z _ {m=cyc.}$                            | Cycle speed                                                     | 14.5                               | MODE_SPEEDS                    | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $Z _ {m=bus}$                             | Bus speed                                                       | 15                                 | MODE_SPEEDS                    | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $Z _ {m=car}$                             | Car speed                                                       | 21                                 | MODE_SPEEDS                    | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $Z _ {m=mot.}$                            | Motorbike speed                                                 | 25                                 | MODE_SPEEDS                    | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| \$\\epsilon \$                            | Walk time to bus                                                | 5 or Lnorm(log (5),0.18)           | BUS\_ WALK_TIME                | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\mu_{m=HGV}$ \|                          | Truck distance relative to car \|                               | 0.21 or Beta(3,10)                 | TRUCK_TO\_ CAR_RATIO \| S      | etting         |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\mu_{m=bus}$ \|                          | Bus driver distance relative to bus passenger distance \|       | 1/45 or Bet a(20,600) \|           | BUS\_ TO_PASSEN GER_RATIO      | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\mu_ {m=mot.}$                           | Motorcycle fleet distance relative to total motorcycle distance | 37/68 or Beta(37/3+ 1,31/3+1) \|   | FLEET_TO_MOTORCYCLE_RATIO      | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $T_j$                                     | Trip-level data                                                 | Constant                           | SYNTHETIC_TRIPS                | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| *Injury model values*                     |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $I _ {a,g,m_{\                            | Fatalities table                                                | Constant                           | INJURY_TABLE                   | Setting        |
| text{vi c} },m_{\text{str}}}$             |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\sigma$                                  | Injury reporting rate                                           | 1 or Beta(8,3)                     | INJURY_REPORTING_RATE          | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\omega$                                  | Sum of distance exponents                                       | 1.9 or Lnorm(l og(1.9),l og(1.03)) | SIN_EXPONENT_SUM               | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\psi$                                    | Casualty fraction of $\omega$                                   | 0.5 or Beta(20,20) \|              | CASUALTY_EXPONENT_FRACTION     | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| -   Pollution model                       |                                                                 |                                    |                                |                |
|                                           |                                                                 |                                    |                                |                |
|     values\*                              |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\eta$                                    | Background PM2.5 concentration                                  | 50 or Lnorm(log( 50),0.18) \|      | PM_CONC_BASE \|                | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\zeta$                                   | Fraction of measured PM2.5 concentration due to road transport  | 0.225 or Beta(5,20) \|             | PM_TRANS_SHARE \|              | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\                                        | Vehicle emission inventory                                      | Constant or Dirichlet              | EMISSION_INVENTORY             | Setting        |
| g amma_{m}$                               |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $P$                                       | Vehicle emission confidence                                     | 1                                  | EM ISSION_INVENTORY_CONFIDENCE | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| -   Pollution & health model              |                                                                 |                                    |                                |                |
|                                           |                                                                 |                                    |                                |                |
|     values\*                              |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $C_1$                                     | Base-level inhalation rate                                      | 1                                  | BAS E_LEVEL_I NHALATION \_RATE | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $C_2$                                     | Exposure ratio window closed                                    | 0.5                                | CLOSED_WI NDOW_PM\_ RATIO      | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $C_3$                                     | Proportion of travel with closed windows                        | 0.5                                | C LOSED_WIN DOW_RATIO          | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $C_4$                                     | Parameter for on-road PM2.5 level                               | 3.216                              | ROAD\_ RATIO_MAX               | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $C_5$                                     | Parameter for on-road PM2.5 level                               | 0.379                              | ROAD_RA TIO_SLOPE              | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $C_6$                                     | Exposure ratio for underground                                  | 0.8                                | SUBWAY \_PM_RATIO              | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| \$ H \_{a,h,w}\$                          | AP dose- -res ponse--curve parameters                           | Constant                           | DR_AP                          | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\                                        | Quantiles for AP RR functions                                   | 0.5 or Unif(0,1)                   | AP_DOSE_R ESPONSE\_ QUANTILE   | Model          |
| xi_{h,w}$                                 |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| *Physical activity model values*          |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\chi$                                    | Day-to-week travel scalar                                       | 7 or 7 $\times$Beta(20,3)          | DAY_TO_WEEK \_TRAVEL_SCALAR    | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $Y_i$                                     | Individual n on-transport MMETs                                 | Constant                           | PA_SET\$ work_ltpa \_marg_met  | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\theta$                                  | Background PA scalar                                            | 1 or Lnorm(0,0.18)                 | BACKGROUND\_ PA_SCALAR         | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\                                        | Confidence in PA survey                                         | 1                                  | BACKGROUND_PA_C ONFIDENCE      | Setting        |
| tilde{Y}$                                 |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\tilde {\theta}$                         | Background PA zeros quantile                                    | 0.5 or Unif(0,1)                   | BACKGROUND \_PA_ZEROS          | Setting        |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\lambda_{m=\text{cyc.}}$                 | Cycling MET surplus                                             | 4.63 or Lnorm(log (4.63),1)        | MMET_CYCLING                   | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\lambda_{m=\text{walk}}$                 | Walking MET surplus                                             | 2.53 or Lnorm(log (2.53),1)        | MMET_WALKING                   | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| $\lambda_{m\not\in\text{wal               | In-vehicle MET surplus                                          | 0                                  |                                | Model          |
| k,cyc.}}$                                 |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| *Physical activity & health model values* |                                                                 |                                    |                                |                |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| \$\\G\_{h,x,z}\$                          | Look-up table for relative risk truncated normal distribution   | Constant                           |                                | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
| \$\phi\_{h}\$                             | Quantiles for PA RR functions                                   | 0.5 or Unif(0,1)                   | PA_DOSE_RESPONSE_QUANTILE      | Model          |
+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+

@@

\[Example of the dose--response workflow for PA and total cancer. Top left: results of the metaanalysis: mean, lower bound and upper bound of the relationship between PA dose and relative risk of disease. Top right: examples of 500 samples from the normal distribution defined by the mean of the top-left panel and the standard deviation defined by (upper bound minus lower bound) divided by 1.96 and truncated at 0. Bottom left: relative risks of individuals with PA doses 5, 25 and 45 are found by mapping their doses onto the median curve. This is used in the "constant" ITHIM use case. Bottom right: relative risks of individuals with PA doses 5, 25 and 45 for three different samples from the distribution shown in the top-right panel. For each sample, each individual is mapped onto the response curve it defines.\]

```{r out.height = "800px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("cancer_PA_dose_response.pdf")
```

@@

The XX uncertain univariate parameters are shown in Figure [3](#distributions){reference-type="ref" reference="distributions"}. Most of the parameters are self explanatory. Cycling and walking MMETs are the number of MMETs per hour when undertaking cycling and walking, and determine also the ventilation rates. Motorcycle distance is the total distance travelled by motorcycles relative to the total distance travelled by cars in the baseline scenario. Non-travel PA, injury reporting rate and NCD burden all act as scalars for the corresponding datasets. Note that the non-travel PA scalar does not affect the $\approx$<!-- -->40% of the population whose non-travel PA is 0.

```{r out.height = "460px", out.width='800px', fig.align='center', fg.cap = 'Distributions of univariate uncertain parameters (1024 samples).', echo=F}
knitr::include_graphics("uncertain_parameters.pdf")
```

(){#distributions width="95%"}

@@

\[Distributions of a Beta-distributed propensity to non-travel PA with varying confidence values in the raw input data. The sampled value gives the fraction of the population that engages in non-travel PA.\]

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("prob_zero_PA.pdf")
```

(){#prob_zero_PA width="50%"}

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
"Distributions of Dirichlet-distributed emission inventories for four modes and varying confidence values. Raw input values were 4, 4, 20, and 60."

knitr::include_graphics("emission_dist.pdf")
```

(){#emission_dist width="65%"}

@@

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F, fig.show='hold', fig.cap="We use the same uniform samples to generate random variables from different distributions. It means that the samples are perfectly correlated."}
knitr::include_graphics("DR_transform.pdf")
```

(){#cdf width="80%"}

@@

# Results {#results}

@@

***YLLs in the scenario minus YLLs in the baseline for Accra***

```{r out.height = "600px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("accra_walk_yll.pdf")
```

@@

\[EVPPI for Accra's "walking" scenario for all causes of YLLs excluding "all cause" and "neoplasm".\]

```{r out.height = "460px", out.width='800px', fig.align='center', fig.cap="EVPPI for Accra",fig.path="accra_walk_evppi.pdf", echo=F}
#knitr::include_graphics("accra_walk_evppi.pdf")
```

# ITHIM with parameters

::: appendix
\[ITHIM-R workflow, with variable parameters in purple. Fixed inputs to the model have a bold, black outline, which is solid for inputs we consider "global" and dashed for inputs we consider "local". There are four global inputs, which will be embedded in ITHIM-R. There are seven local inputs, which users should provide. The graph depicts parent--child relationships, where a child (at the head of an arrow) depends upon all its parents (at the source(s) of the arrow(s)). To aid visualisation, an arrow connected to the outside of a grey box indicates that the arrow connects to each item within the box. *AT: active travel. MET: metabolic equivalent task. pp: per person. AP: air pollution. PA: physical activity. DR: dose--response relationship. GBD: global burden of disease.* NB: $\rho$ only impacts the NCD burden of GBD, not all of it. Missing parameter: motorcycle distance. $\lambda$ is a 2D parameter. $\xi$ is 4D.\]

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("ithim_with_parameters.pdf")
```

(){#ithim_with_parameters width="\\textwidth"}

@@

\[1000 samples from PA dose--response curves\]

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("PA_dose_response.pdf")
```

(){#PADR width="90%"}

+:----------------------------------------------------------------------------------:+:---------------------------------------------------------------------------------:+
| ![100 samples from AP dose--response curves](cvd_ihdDR-sim.pdf){#APDR width="40%"} | \[100 samples from AP dose--response curves\] (cvd_strokeDR-sim.pdf){width="40%"} |
+------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+

+:-----------------------------------------------------------------:+:------------------------------------------------------------:+:------------------------------------------------------------------:+
| ![100 samples from AP dose--response curv es](neo_lungDR-sim.pdf) | ![100 samples from AP dose--response c urves](lriDR-sim.pdf) | ![100 samples from AP dose--response curve s](resp_copdDR-sim.pdf) |
+-------------------------------------------------------------------+--------------------------------------------------------------+--------------------------------------------------------------------+

::: landscape
# Tabulated ITHIM equations {#tableequations}

::: center
::: ThreePartTable
::: TableNotes
CDF: cumulative distribution function. We use generative distributions to generate random numbers so that samples are correlated. For $t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$ with CDF CDF$_F$. (I'd like to find an improved notation for this.)
:::

@@

+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| Name                                                                                                                                                         | Description                                                                                    | Label                                        |
+:=============================================================================================================================================================+:===============================================================================================+:=============================================+
| *Scenario generation*                                                                                                                                        |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| ...                                                                                                                                                          |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| *Travel calculations*                                                                                                                                        |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk} \}\, \forall\, T_j:m(j)=\text{bus}$                                                               | New walk trips to bus                                                                          | walk *to_bus_and* combine_scen               |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
